---
title: "Methodology"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Methodology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

For instructions on using this package, please see the<a href = "Manual.html">package Manual</a>.

## Rationale

- Wash, fsl and capacity gaps can each drive up the msni by themselved (i.e. if Wash is 3, the msni will be at least 3 no matter what)  
    - capacity gaps is included to take into account that households may have low living standard gaps at the expense of extreme coping strategies
- Health, protection and shelter drive up the msni in combinations of two.
- education: Education living standard gaps contribute to the msni, but do not produce "extreme" (4) situations if needs are already high in for example food security and wash   
- impact: this is used as a "contributing factor"; with high impact scores, any sectoral sub index can drive up the msni  

## Implementation

### Description

The calculation can be described as follows:

For each household individually...

  1. set `msni` to the largest value of `health`, `protection` and `shelter`
  2. set `msni` to the `impact` score, if `impact` is lower than the previous `msni`. 
  3. set `msni` to the largets value of `msni`, `wash`, `fsl`, `capacity gaps` if that is larger than the previous `msni`
  4. set `msni` to `2` _if and only if_
      1. `msni` is currently `1`
      2. _and_ one of health, protection, shelter and education is larger or equal `3`
  5. set `msni` to `NA` if *any* of the input scores was `NA`
  6. done!

### Code

Here is the actual code.

```{r, eval = FALSE}

#  1. set `msni` to the largest value of `health`, `protection` and `shelter`
msni <- pmax(health_lsg,protection_lsg,shelter_lsg)

#  2. set `msni` to the `impact` score, if `impact` is lower than the previous `msni`. 
msni <- pmin(msni, impact)

#  3. set `msni` to the largets value of `msni`, `wash`, `fsl`, `capacity gaps` if that is larger than the previous `msni`
largest_lsg_combination_hps <- pmax(
  pmin(health_lsg,protection_lsg),
  pmin(health_lsg, shelter_lsg),
  pmin(protection_lsg,shelter_lsg)
)

msni <- pmax(msni, largest_lsg_combination_hps)

# If higher than current, replace with highest from wash, FSL or capacity gap
msni <- pmax(msni, wash_lsg,fsl_lsg,capacity_gaps)


#4.A is `msni` is currently `1` _and_ one of health, protection, shelter and education is larger than `3` ?
replacement_condition <- (msni == 1 & pmax(health_lsg,
                                           protection_lsg,
                                           shelter_lsg,
                                           education_lsg) >= 3)

# 4.B set `msni` to `2` where condition above applies
msni[replacement_condition] <- 2

```

Dictionary for non-programmers:

- ` <- ` means "resplace" or "assign" the value to the name
- ` pmin() ` the smallest of the values in parenthesis (per record)
- ` pmax() ` the largest of the values in parenthesis (per record)
- `variable[condition] <-` replace / assign only if/where condition is met



